Viscous Erosion in Porous Media

### 2018-02-26 docs/methods/erosion_methods.tex
NM

Erosion is a fluid-mechanical process responsible for a variety
of phenomena in geophysical and other settings.  For example,
wind erosion causes sand dunes to be  longer on the windward
side~\cite{han1969}, while stream-bed erosion is responsible for branching structures
of rivers~\cite{coh-dev-sey-yi-szy-rot2015}.  Erosion is
also important in biophysical applications, for example plaque erosion in arteries
is key to understanding heart disease~\cite{sha2002,
gro-gij-van-fer-hat-van-yua-wen2007}, and the separation of biofilms can
be explained by erosion~\cite{pic-van-hei2000}. 
Much recent work has been devoted to understanding fundamental aspects of the fluid-structure interaction that occurs during the erosion \cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, mit-spa2016, hewett2017evolution, moore2017riemann, lachaussee2018competitive, lopez2018cfd}, and related processes like dissolution or melting \cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, yuan2017dynamics, claudin2017dissolution, hewett2017pear}. All of these processes involve a feedback process between how boundaries move and how flow evolves and thus can be conceptually linked  as `sculpting with flow' \cite{ristroph2018sculpting}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary evolution in the {\thL} framework} 
\label{sec:thetaL}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}
\label{s:results}
With the description of the methods complete, we now present numerical
results on bodies eroding in Stokes flow. We first discuss the case of a
single body, as certain features can be predicted analytically and thus
used for validation. We then discuss the erosion of multiple bodies and
highlight important differences that arise compared to the single-body
case.

\subsection{Shape evolution of a single eroding body}

@@

\subsection{The limiting shape}

@@

\subsection{Convergence tests}

As a final validation, we now aim to verify the second-order temporal convergence rate of our method. To this end, we run the same single-body simulation from Figs.~\ref{01bodseq} and \ref{shrink_intface} (except with $N_\iin=128$) to a stopping time of $t_s = 1E-2$. This stopping time is 56\% of the time it takes for the body to vanish completely, so that the body has eroded substantially, but is not so small that accuracy is lost. We assess error by measuring the difference between simulations of successive temporal resolution (i.e.~we verify Cauchy convergence). In particular, we measure the $L^2$-difference of the body shapes obtained at the stopping time, $t_s$, while successively halving the timestep $\Delta t$. Table \ref{convtab} shows the $L^2$-errors between successive simulations, along with the order of accuracy as obtained by Richardson-extrapolation-type calculations. The table confirms that our method to converges with second-order accuracy.

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{Convergence test: We run erosion simulations to a stopping time of $t_s =$ 1E-2 (56\% of the vanishing time) and measure the $L^2$ difference in shapes obtained with successive temporal resolution. The tests confirm the method to be second-order accurate in time.
}
\vspace{0.3 pc}
\label{convtab}
\begin{tabular}{l l l l}
\hline
\hspace{0.0pc} $\Delta t/t_s$
\hspace{0.5pc} & error 
\hspace{0.5pc} & order
\hspace{0.5pc} & CPU time \\
\hline
%
1/50		& 1.68E-5		& --		& 55 secs		\\
1/100	& 4.84E-6		& 1.79	& 1.8 mins		\\
1/200	& 1.30E-6		& 1.90	&	\\
1/400	& 3.36E-7		& 1.95	&	\\
1/800	& 8.54E-8		& 1.98	&	\\
1/1600	& 2.15E-8		& 1.99	&	\\
1/3200	& 5.40E-9		& 1.99	&	\\
1/6400	& --			& --		&	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Parameters in this test are: 01circ256, nits = 7, dt0 = 1e-4, tfin = 1e-2
% epsfac = 10, sigfac = 10, fixarea = 0, fixprdop = 0

\todo[inline]{Could put CPU time in table too}

@@

Runs to do
\begin{itemize}
  \item Single body
  \begin{itemize}
    \item Aspect ratio and opening angle
    \item Drag and resistivity
    \item Anisotropy
    \item Vorticity
  \end{itemize}
  \item $\bigO(5)$ bodies
  \begin{itemize}
    \item Drag and resistivity
    \item Anisotropy
    \item Long channels between flat faces
    \item Vorticity
  \end{itemize}
  \item $\bigO(50)$ bodies
  \begin{itemize}
    \item Drag and resistivity
    \item Anisotropy
    \item Vorticity
  \end{itemize}
\end{itemize}

>>>>

Erosion is a fluid-mechanical process responsible for a variety
of phenomena in geophysical and other settings.  For example,
wind erosion causes sand dunes to be  longer on the windward
side~\cite{han1969}, while stream-bed erosion is responsible for branching structures
of rivers~\cite{coh-dev-sey-yi-szy-rot2015}.  Erosion is
also important in biophysical applications, for example plaque erosion in arteries
is key to understanding heart disease~\cite{sha2002,
gro-gij-van-fer-hat-van-yua-wen2007}, and the separation of biofilms can
be explained by erosion~\cite{pic-van-hei2000}. 
Much recent work has been devoted to understanding fundamental aspects of the fluid-structure interaction that occurs during the erosion \cite{ris-moo-chi-she-zha2012, moo-ris-chi-zha-she2013, mit-spa2016, hewett2017evolution, moore2017riemann, lachaussee2018competitive, lopez2018cfd}, and related processes like dissolution or melting \cite{Huang2015, kondratiuk2015steadily, rycroft2016asymmetric, cohen2016erosion, yuan2017dynamics, claudin2017dissolution, hewett2017pear}. All of these processes involve a feedback process between how boundaries move and how flow evolves and thus can be conceptually linked  as `sculpting with flow' \cite{ristroph2018sculpting}.
% Later, we can take out a lot of these references, especially ones that are cited elsewhere in the paper. For example, Mitchell and Spagnolie and several of my papers are cited elsewhere.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Boundary evolution in the {\thL} framework and regularization} 
\label{sec:thetaL}

@@

%%-----------------------------------------------------------------%%
% RESULTS SECTION
%%-----------------------------------------------------------------%%
\section{Results}
\label{s:results}
With the description of the methods complete, we now present numerical
results on bodies eroding in Stokes flow. We first discuss the case of a
single body, as certain features can be predicted analytically and thus
used for validation. We then discuss the erosion of multiple bodies and
highlight important differences that arise compared to the single-body
case.

\subsection{A single body eroding in Stokes flow}

@@

\subsection{Local analysis for the limiting shape}

@@

\subsection{Convergence test}

As a final validation, we now aim to verify that our method converges with second-order accuracy in time. To this end, we run the same single-body simulation from Figs.~\ref{01bodseq} and \ref{shrink_intface} (except with a coarser $N_\iin=256$) to a stopping time of $t_s = 1E-2$. This stopping time is 56\% of the time it takes the body to vanish, and was chosen so that the body will have eroded substantially but is not so small that accuracy is lost. We assess error by measuring the difference between simulations of successive resolutions (i.e.~we verify Cauchy convergence). In particular, we measure the $L^2$-difference of the body shapes obtained at the stopping time. At each stage we halve the timestep $\Delta t$. Table \ref{convtab} shows the $L^2$-errors between successive simulations, along with the order of accuracy as obtained by Richardson-extrapolation-type calculations. The table confirms that our method to converges with second-order accuracy. In the table, we also show the CPU time of the simulation run on a single processor (though the algorithm is easily parallelized), showing a linear increase in computational time.

% Table
%^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
\begin{table}%[htbp]
\begin{center}
\caption{Convergence test: We run erosion simulations to a stopping time of $t_s =$ 1E-2 (56\% of the vanishing time) and measure the $L^2$ difference in shapes obtained with successive temporal resolution. The tests confirm the method to be second-order accurate in time.
}
\vspace{0.3 pc}
\label{convtab}
\begin{tabular}{l l l l}
\hline
\hspace{0.0pc} $\Delta t/t_s$
\hspace{0.5pc} & error 
\hspace{0.5pc} & order
\hspace{0.5pc} & CPU time \\
\hline
%
1/50		& 1.68E-5		& --		& 55 secs		\\
1/100	& 4.84E-6		& 1.79	& 1.8 mins		\\
1/200	& 1.30E-6		& 1.90	& 3.5 mins		\\
1/400	& 3.36E-7		& 1.95	& 7.2 mins		\\
1/800	& 8.54E-8		& 1.98	& 15 mins		\\
1/1600	& 2.15E-8		& 1.99	& 30 mins (?)	\\
1/3200	& 5.40E-9		& 1.99	& 60 mins (?)	\\
1/6400	& --			& --		& 2.0 hours(?)	\\
%
\hline
\end{tabular}
\end{center}
\end{table}
 %^^^^^^^^^^^^^^^^^^^^^^^^^^^^^^%
% Parameters in this test are: 01circ256, nits = 7, dt0 = 1e-4, tfin = 1e-2
% epsfac = 10, sigfac = 10, fixarea = 0, fixprdop = 0

\subsection{Multiple-body simulations}
With the numerical method validated, we are now ready to simulate the erosion of multiple bodies.

Possible list:
\begin{itemize}
\item Show multiple body simulation with fixed flow rate
\item Then simulation with fixed pressure drop
\end{itemize}

@@

Runs to do
\begin{itemize}
  \item Single body
  \begin{itemize}
    \item Drag and resistivity
  \end{itemize}
  \item $\bigO(5)$ bodies
  \begin{itemize}
    \item Drag and resistivity
    \item Anisotropy
    \item Long channels between flat faces
    \item Vorticity
  \end{itemize}
  \item $\bigO(50)$ bodies
  \begin{itemize}
    \item Drag and resistivity
    \item Anisotropy
    \item Vorticity
  \end{itemize}
\end{itemize}
